---
title       : Data Visualization for Education
subtitle    : SDP Fellow Workshop January 2013
author      : Jared Knowles
job         : Policy Research Advisor, Wisconsin DPI
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : prettify  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : standalone # {standalone, draft}
---

## Objectives

1. Review data visualization principles
2. Look at applications in education data
3. Challenges in an LEA/SEA
4. Best practices and advice
5. What tools to use

```{r knitrsetup,echo=FALSE,eval=TRUE,results="hide"}
library(knitr)
opts_chunk$set(fig.width=8,fig.height=5.5,message=FALSE,error=FALSE,warning=FALSE,echo=TRUE,dev="svg",comment=NA)
library(eeptools)
library(ggplot2)
library(ggthemes)
```

--- plot

### Example

```{r plot, fig.height=5.5, fig.width=8.5,fig.align="center",out.width="700px",out.height="500px"}
qplot(hp,mpg,data=mtcars)+theme_dpi()
```

--- plot

### Example

```{r plot1,out.width="300px",out.height="200px",echo=FALSE}
qplot(hp,mpg,data=mtcars)+theme_dpi()
qplot(hp,data=mtcars)+theme_dpi()
qplot(drat,data=mtcars,geom="density")+theme_dpi()
qplot(factor(cyl),disp,data=mtcars,geom="boxplot")+theme_dpi()
qplot(factor(vs),mpg,data=mtcars,geom="violin")+theme_dpi()
qplot(hp,qsec,data=mtcars,geom="point")+theme_dpi()+geom_smooth()
```

--- &twocol w1:40% w2:60%

## My Header

*** left

- What to do?
- When the project?
- Is too big?

*** right

- Don't panic
- Relax
- Breathe

--- bg:url(img/badinfographic.png) bg-repeat: no-repeat

### Arbitrary image

--- quote

<p><q> Think like a <span class = 'red'>map</span>. Data density and easy interpretability.</q></p>

---

### Some tips

>- Focus on the content
>- Use best practices
>- Understand the limitations
>- Experiment and iterate!

---

### Some tips

>- Have a properly chosen format and design 
>- Use words, numbers and drawing together
>- Reflect a balance, a proportion, relevant scale
>- Display an accessible complexity of details
>- Have a narrative quality, tell a story
>- Avoid content-free decoration (Tufte's proverbial chartjunk)
>- Draw in a professional manner with an eye on the technical details
>- Remember the map

--- plot

### Themes convey brands


```{r plot2, fig.height=5.5, fig.width=8.5,fig.align="center",out.width="750px",out.height="480px"}
qplot(hp,mpg,data=mtcars)+theme_economist()
```

--- plot

### They Also Communicate


```{r plot3, fig.height=5.5, fig.width=8.5,fig.align="center",out.width="750px",out.height="480px"}
qplot(hp,mpg,data=mtcars)+theme_tufte()
```


--- plot

### They Also Can Confound


```{r plot4, fig.height=5.5, fig.width=8.5,fig.align="center",out.width="750px",out.height="480px"}
qplot(hp,mpg,data=mtcars,color=factor(cyl))+theme_excel2003()+scale_color_excel2003()
```


--- plot

### So Choose Wisely


```{r plot5, fig.height=5.5, fig.width=8.5,fig.align="center",out.width="750px",out.height="480px"}
qplot(hp,mpg,data=mtcars,color=factor(cyl))+theme_stata()
```

---

---

### Ugly graphic

<script>
    function init() {

      var canvas = document.getElementById("canvas");
      var ctx = canvas.getContext("2d");

      draw(ctx);
    }

    function draw(ctx) {

      var gradient;

      // layer1/Path
      ctx.save();
      ctx.beginPath();
      ctx.moveTo(0.5, 0.0);
      ctx.lineTo(0.5, 423.0);
      ctx.stroke();

      // layer1/Path
      ctx.beginPath();
      ctx.moveTo(0.0, 423.5);
      ctx.lineTo(477.0, 423.5);
      ctx.stroke();

      // layer1/This is silly
      ctx.font = "36.0px 'Nyala'";
      ctx.fillText("This gradient is silly", 62.1, 93.3);

      // layer1/Path
      ctx.beginPath();
      ctx.moveTo(472.0, 317.0);
      ctx.lineTo(62.0, 317.0);
      ctx.lineTo(62.0, 175.0);
      ctx.lineTo(472.0, 175.0);
      ctx.lineTo(472.0, 317.0);
      ctx.closePath();
      gradient = ctx.createRadialGradient(267.0, 246.0, 0.0, 267.0, 246.0, 153.4);
      gradient.addColorStop(0.50, "rgb(255, 255, 255)");
      gradient.addColorStop(0.65, "rgb(255, 169, 171)");
      gradient.addColorStop(1.00, "rgb(255, 83, 88)");
      ctx.fillStyle = gradient;
      ctx.fill();
      ctx.restore();
    }
</script>

<body onload="init()">
   <canvas id="canvas" width="477" height="424"></canvas>
</body>

---

### Backmatter

```{r wrapup,echo=TRUE,eval=TRUE,results="markup"}
print(sessionInfo(),locale=FALSE)
```